Model Search in Relational Domains

### 2016-10-21 learning.tex

\section{Abstract Conflict Analysis}
A conflict analysis in a SAT solver can be seen as abductive 
reasoning which underapproximate a set of models that do not satisfy a 
formula~\cite{sas12,dhk2013-popl}.  
\Omit {
where a singleton assignment $c$ is replaced by a partial 
assignment that is sufficient to infer $c$.  Thus, a conflict 
analysis is a generalization of an under-approximation of a set of 
countermodels, where a countermodel is a 
set of structures that do not satisfy a formula 
}
Since relational abstract domains have more complex structure than a partial 
assignments domain or interval domain, so finding a generalized reason for 
conflict through UIP computation~\cite{uip} is non-trivial.

\paragraph {\textbf{Abstract Conflict Graph}}
The first-UIP algorithm in SAT solver works on a data structure called 
{\em implication graph} that computes a cut that suffices to produce a 
conflict.  The trail in ACDCL implicitly encodes a graph structure called 
{\em abstract conflict graph}, which records the dependencies between 
deductions made during abstract model search phase.  Nodes in the graph 
represent elements of the trail $\mathcal{T}$.  Nodes can be either 
{\em decision} node or {\em deduction} node. Edges can be extracted from
the transformers in the reason trail.  Incoming arrows to node $n$ 
indicate that the predecessors of $n$ are sufficient to deduce $n$.  
A decision node has no incoming edges.  
For example, consider a formula 
$\varphi:= (y=x \; \wedge \; y=y+x \; \wedge \; y \leq 0)$ with current 
abstract valuation $v:= (x-5 \geq 0 \wedge -x+5 \geq 0)$.  
Figure~\ref{conflict} shows a snapshot of an abstract conflict graph that 
stores the deductions obtained from abstract model search using octagon domain.  
The last deduction conflict with the constraint $(y \leq 0)$.

@@

involve meet irreducibles that contains the same variables.  The 

@@

yields clauses that generates new deduction after backtracking.  

@@

For our example, there exists multiple incomaparable reasons for conflict,
marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  

@@

\paragraph{\textbf{Learning in Relational Lattice}}
Learning in propositional solver yields a asserting clause~\cite{cdcl} 
that express the negation of the conflict reasons.  However, we model 

@@

{\em abstract unit transformer (AUnit)}.  We add $AUnit$ to the set of 

@@

of meet irreducibles $C \subseteq A$ of $\varphi$ such that $\bigsqcap C$ does not 
satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is formally defined as follows. 

@@

$AUnit$ returns $\bot$ if $C$ is conflicting following rule (1).  Rule (2) and
(3) of $AUnit$ infers a valid meet irreducible, which implies that $C$ is unit. 

@@

domains.  For example, let $C = \langle x+z \geq 10,  w \geq 0 \rangle$ 

@@

x+z \leq 15, w \geq 1, w \leq 1 \rangle$.  Applying rule (3) of $AUnit_C$, we get a 

@@

$a''= \langle w \geq 1, w \leq 1 \rangle \sqsubseteq (w \geq 0)$.  Then, $AUnit_C(a) = (x+z < 10)$, 

@@

and let $a = \langle x \geq 3, x \leq 4, y \geq 5, y \leq 6 \rangle$, then $AUnit_C(a) = \bot$ using

@@

then $AUnit_C(a') = (y > 7)$ using rule (2), since $a' \sqsubseteq (x \geq 2)$
and $a' \sqsubseteq (x \leq 5)$.  However, if none of the above rule holds true, $AUnit$ 

@@

\paragraph {\textbf{Backjumping}}
A backjumping procedure undo all the assignments in the trail up to 
a decision level that restores the solver to a consistent state 
(non-conflicting).  The backjumping level is defined by the literal 
of the conflict clause that is closest to the root (decision level 0) 
where the conflict clause is still unit. If a conflict clause is 
globally unit, then the backjumping level is the root of the search tree.

>>>>

\section{Abstract Conflict Analysis}
Conflict analysis in a SAT solver can be seen as abductive 
reasoning which underapproximates a set of models that do not satisfy a 
formula~\cite{sas12,dhk2013-popl}.  
\Omit {
where a singleton assignment $c$ is replaced by a partial 
assignment that is sufficient to infer $c$.  Thus, a conflict 
analysis is a generalization of an under-approximation of a set of 
countermodels, where a countermodel is a 
set of structures that do not satisfy a formula 
}
Since relational abstract domains have more complex structure than the partial 
assignments domain or interval domain, so finding a generalized reason for 
conflict through UIP computation~\cite{uip} is non-trivial.

\paragraph {\textbf{Abstract Conflict Graph}}
The first-UIP algorithm in SAT solvers works on a data structure called 
{\em implication graph} and computes a cut that suffices to produce a 
conflict.  The trail in ACDCL implicitly encodes a graph structure called 
{\em abstract conflict graph}, which records the dependencies between 
deductions made during the abstract model search phase.  Nodes in the graph 
represent elements of the trail $\mathcal{T}$.  A node can be either a
{\em decision} node or a {\em deduction} node. Edges can be extracted from
the transformers in the reason trail.  Incoming arrows to node $n$ 
indicate that the predecessors of $n$ are sufficient to deduce $n$.  
A decision node has no incoming edges.  
For example, consider a formula 
$\varphi:= (y=x \; \wedge \; y=y+x \; \wedge \; y \leq 0)$ with current 
abstract valuation $v:= (x-5 \geq 0 \wedge -x+5 \geq 0)$.  
Figure~\ref{conflict} shows a snapshot of an abstract conflict graph that 
stores the deductions obtained from abstract model search using the octagon domain.  
The last deduction conflicts with the constraint $(y \leq 0)$.

@@

involve meet irreducibles that contain the same variables.  The 

@@

yields clauses that generate new deductions after backtracking.  

@@

For our example, there exist multiple incomaparable reasons for conflict,
marked as {\em cut0, cut1}, in Figure~\ref{conflict}.  Here, cut0 is the first UIP.  

@@

\paragraph{\textbf{Learning in relational domains}}
Learning in a propositional solver yields an asserting clause~\cite{cdcl} 
that expresses the negation of the conflict reasons.  However, we model 

@@

{\em abstract unit transformer (AUnit)}.  We add $\mathit{AUnit}$ to the set of 

@@

of meet irreducibles $C \subseteq A$ of $\varphi$ \pscmt{???} such that $\bigsqcap C$ does not 
satisfy $\varphi$, $\mathit{AUnit}_C: A \rightarrow A$ is formally defined as follows. 

@@

$\mathit{AUnit}$ returns $\bot$ if $C$ is conflicting following rule (1).  Rule (2) and
(3) of $\mathit{AUnit}$ infers a valid meet irreducible, which implies that $C$ is unit. 

@@

domains \pscmt{???}.  For example, let $C = \langle x+z \geq 10,  w \geq 0 \rangle$ 

@@

x+z \leq 15, w \geq 1, w \leq 1 \rangle$.  Applying rule (3) of $\mathit{AUnit}_C$, we get a 

@@

$a''= \langle w \geq 1, w \leq 1 \rangle \sqsubseteq (w \geq 0)$.  Then, $\mathit{AUnit}_C(a) = (x+z < 10)$, 

@@

and let $a = \langle x \geq 3, x \leq 4, y \geq 5, y \leq 6 \rangle$, then $\mathit{AUnit}_C(a) = \bot$ using

@@

then $\mathit{AUnit}_C(a') = (y > 7)$ using rule (2), since $a' \sqsubseteq (x \geq 2)$
and $a' \sqsubseteq (x \leq 5)$.  However, if none of the above rule holds true, $\mathit{AUnit}$ 

@@

\paragraph {\textbf{Backjumping}}
A backjumping procedure undoes all the assignments in the trail up to 
a decision level \pscmt{explained?} that restores the solver to a consistent state 
(non-conflicting).  The backjumping level is defined by the literal 
of the conflict clause that is closest to the root (decision level 0) 
where the conflict clause is still unit. If a conflict clause is 
globally unit, then the backjumping level is the root of the search tree.
